---
title: "p8105_hw3_jae2186"
author: "jae2186 (Jennifer Estrada)"
date: "10/20/2021"
output:
  github_document
---

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(p8105.datasets)
library(ggridges)
library(lubridate)

knitr::opts_chunk$set(
	fig.asp = 0.75,
	fig.width = 6,
	message = FALSE,
	warning = FALSE,
	out.width = "90%"
)

theme_set(theme_bw() + theme(plot.caption= element_text(size = 6.5)))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  knitr.kable.NA = ""
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Problem 1 - Instacart Data

```{r ic_description}

instacart_df <- force(instacart)

insta_ohod <-
  instacart_df %>% 
  group_by(order_hour_of_day) %>% 
  summarize(n_obs = n(), n_orders = n_distinct(order_id)) %>% 
  arrange(desc(n_orders)) %>% 
  mutate(order_hour_of_day = hms::as_hms(3600 * order_hour_of_day))

insta_dow <-
  instacart_df %>% 
  group_by(order_dow) %>% 
  summarize(n_obs = n(), n_orders = n_distinct(order_id)) %>% 
  arrange(desc(n_orders)) %>% 
  mutate(order_dow = wday(order_dow + 1, label = TRUE, abbr = FALSE))

# The above code for `insta_dow` works on the assumption that weekdays were coded to start on Sunday, which was coded as '0'.

insta_dbo <-
  instacart_df %>% 
  group_by(days_since_prior_order) %>% 
  summarize(n_obs = n(), n_orders = n_distinct(order_id)) %>% 
  arrange(desc(n_orders))

insta_prod_name <- 
  instacart_df %>% 
  select(product_name) %>% 
  count(product_name, name = "prod_count") %>% 
  arrange(desc(prod_count))

insta_aisle <-
  instacart_df %>% 
  select(aisle) %>% 
  count(aisle, name = "aisle_count") %>% 
  arrange(desc(aisle_count))

```

This Instacart dataset has 15 variables describing `r instacart_df %>% group_by(order_id) %>% summarize(n_obs = n()) %>% count()` individual orders which included `r length(instacart_df$order_id)` items ordered; each order was made by a unique Instacart user during 2017. The data include aspects regarding the timing the order including the time of day at which and day of the week on which orders were placed. For instance: 

* Orders are typically made in the middle of the day, which is shown by the fact that the top 3 hour-long periods for orders were: `r insta_ohod[1,1]` (or 2 pm), `r insta_ohod[2,1]` (or 3 pm), and `r insta_ohod[3,1]` (or 1 pm). (Nearly half-a-day apart from the hour at which the fewest orders were placed: `r insta_ohod[24,1]`, or 4 am.)

* Orders were placed most frequently on `r insta_dow[1,1]` while the fewest orders were placed on `r insta_dow[7,1]`.

The data also consider characteristics of the order that were unique to the user (i.e. which number order this was for the shopper and the number of days since their last order). The data note that:

* Most the of the orders were placed `r insta_dbo[1,1]` days (or about 1 month) following the previous order. This is in stark contrast to that fact that the second most frequent days between orders was `r insta_dbo[2,1]` days, or 1 week.

Lastly, the dataset addresses details about the items in the cart, including the order in which items are put into the virtual carts, the name of the item, and the location of the item in the grocery store (i.e. aisle and department). Here are some quick notes on this dataset:

* The data note that there are `r instacart_df %>% group_by(product_id) %>% summarize(n_obs = n()) %>% count()` different products ordered, with "`r insta_prod_name[1,1]`" being the most common product ordered while items from the `r insta_aisle[1,1]` and `r insta_aisle[2,1]` aisles ordered the most (with a difference of `r insta_aisle[1,2]-insta_aisle[2,2]` orders compared to the following aisle of `r insta_aisle[3,1]` which had `r insta_aisle[2,2]-insta_aisle[3,2]` fewer orders compared to `r insta_aisle[2,1]`).

* When answering: <i>how many aisles are there, and which aisles are the most items ordered from? </i>
  * <b>There are `r max(instacart$aisle_id)` aisles and, as previously noted, the top three aisles from which items were ordered included: 
    * `r insta_aisle[1,1]` (with `r insta_aisle[1,2]` orders), 
    * `r insta_aisle[2,1]` (with `r insta_aisle[2,2]` orders), and 
    * `r insta_aisle[3,1]` (with `r insta_aisle[3,2]` orders). </b>
  * Of note, these are all produce aisles. 
  
The following includes a plot that shows the number of items ordered in each aisle, limiting this to aisles with >10000 items ordered.

```{r ic_num_items}

instacart_df %>% 
  count(aisle_id) %>% 
  filter(n > 10000) %>% 
  ggplot(aes(x = aisle_id, y = n)) + 
  geom_point() + 
  scale_y_continuous(trans = "log10", labels = scales::label_number()) +
  scale_x_continuous(breaks = seq (0, 175, by = 25)) +
  annotation_logticks(sides = "l") +
  labs(
    title = "Number of items ordered per aisle", 
    subtitle = "Minimum of 10,000 items ordered", 
    x = "Aisle Number",
    y = "Items Ordered (log10 scale)",
    caption = "Data from “The Instacart Online Grocery Shopping Dataset 2017”, https://www.instacart.com/datasets/grocery-shopping-2017 on 10/20/2021.")

```

Below is a table showing the three most popular items in each of the aisles categorized as “baking ingredients”, “dog food care”, and “packaged vegetables fruits”, with the number of times each item is ordered. 

```{r ic_pop_items, message=FALSE, warning=FALSE}

pop_items <-
  instacart_df %>% 
  select(aisle, product_name) %>% 
  filter(aisle == "packaged vegetables fruits"|
           aisle == "baking ingredients" |  
           aisle == "dog food care"
                    ) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_obs = n()) %>% 
  pivot_wider(names_from = aisle, values_from = n_obs) %>% 
  filter(min_rank(desc(`packaged vegetables fruits`)) <4 |
           min_rank(desc(`baking ingredients`)) <4 |
           min_rank(desc(`dog food care`)) <4) %>% 
  knitr::kable(col.names = c("Product Name", "Baking Ingredients", 
                             "Dog Food Care", 
                             "Packaged Vegetables Fruits"),
               align = "lccc",
               caption = "3 Most Popular Products Ordered via Instacart for 3 Aisles")

pop_items
```

Below is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r ic_mean_hr}

mean_hod <- 
  instacart_df %>% 
  select(product_name, order_dow, order_hour_of_day) %>%
  filter(product_name == "Pink Lady Apples" | 
           product_name == "Coffee Ice Cream") %>% 
  mutate(order_dow = wday(order_dow + 1, label = TRUE, abbr = FALSE)) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_hod = mean(order_hour_of_day)
  ) %>% 
  rename("Product Name" = product_name) %>% 
  pivot_wider(names_from = order_dow, values_from = mean_hod) %>% 
  knitr::kable(align = "lccccccc", digits = 1,
               caption = "Mean hour of the day at which 2 items are ordered each day of the week via Instacart")

mean_hod
```


## Problem 2 - BRFSS SMART 2010 Data

This problem uses data made available on data.gov from the Behavioral Risk Factors Surveillence System (BRFSS) for Selected Metropolitan Area Risk Trends (SMART) from the years 2002-2010.

```{r brfss_data}
brfss_df <- force(brfss_smart2010) 

brfss_clean <- 
  janitor::clean_names(brfss_df) %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Excellent" |
           response == "Very good" |
           response == "Good" |
           response == "Fair" |
           response == "Poor"
           ) %>% 
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent")))

brfss_02 <-
  filter(brfss_clean, year == 2002) %>% 
  select(contains("location")) %>% 
  group_by(locationabbr, locationdesc) %>% 
  summarize(n_obs = n()) %>% 
  count(locationabbr) %>% 
  arrange(desc(n)) %>% 
  filter(n > 6)

brfss_10 <-
  filter(brfss_clean, year == 2010) %>% 
  select(contains("location")) %>% 
  group_by(locationabbr, locationdesc) %>% 
  summarize(n_obs = n()) %>% 
  count(locationabbr) %>% 
  arrange(desc(n)) %>% 
  filter(n > 6)
```

Given the above data set it was noted that:

* In 2002 there were `r length(brfss_02$locationabbr)` states observed at 7 or more locations including: `r brfss_02$locationabbr`.

* In 2010 there were `r length(brfss_10$locationabbr)` states observed at 7 or more locations including: `r brfss_10$locationabbr`.

* Between 2002 and 2010 there were `r length(brfss_10$locationabbr)- length(brfss_02$locationabbr)` more states with at least 7 observation sites; not all those originally with 7 or more sites in 2002 maintained these observation locations as CT dropped down to 5 locations.

The following 2 plots further describe aspects of the data reported:

```{r brfss_spaghetti}
brfss_avg_val <-
  filter(brfss_clean, response == "Excellent") %>% 
  select(year, locationabbr, locationdesc, data_value) %>% 
  group_by(year, locationabbr) %>% 
  mutate(avg_val = mean(data_value)) %>% 
  ggplot(aes(x = year, y = avg_val, color = locationabbr)) +
  geom_line() +
  theme(legend.position = "none") +
  labs(
    title = "Spaghetti plot of BRFSS data for observed US states from 2002-10", 
    subtitle = "Average values of 'Excellent' responses to 'Overall Health' Questions", 
    x = "Year",
    y = "Average Value",
    caption = "Data obatined from publically available 2002-2010 BRFSS SMART County Prevalence land line only data, from data.gov on 10/20/2021.")

brfss_avg_val
```

```{r brfss_NY}
brfss_ny <-
  filter(brfss_clean, locationabbr == "NY" & year == "2006" |
           locationabbr == "NY" &
           year == "2010") %>% 
  select(year, locationdesc, response, data_value) %>% 
  group_by(year, locationdesc) %>% 
  ggplot(aes(x = data_value, y = locationdesc)) + 
  geom_density_ridges(scale = .85) +
  facet_grid(. ~ year) +
  labs(
    title = "BRFSS data for observed NY counties in 2006 and 2010", 
    subtitle = "Density plot of data values to 'Overall Health' Questions", 
    x = "Data Value",
    y = "New York County",
    caption = "Data obatined from publically available 2002-2010 BRFSS SMART County Prevalence land line only data, from data.gov on 10/20/2021.") +
  scale_y_discrete(
    labels = c("Westchester", "Suffolk", "Queens", "New York", 
               "Nassau","Monroe", "Kings", "Erie", "Bronx")
  )
  
brfss_ny
```

## Problem 3 - Accelerometer Data

```{r accel_data}

accel_df <-
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    day = recode(day, `Monday` = "1", `Tuesday` = "2", `Wednesday` = "3", `Thursday` = "4",
                 `Friday` = "5", `Saturday` = "6", `Sunday` = "7"),
    day = parse_number(day),
    week_v_end = ifelse(day <= 5, 1, 0)
    ) %>% 
  relocate(week, day_id, day, week_v_end)

```

The dataset consists of acceleramator activity data over the course of `r max(accel_df$week)` weeks, for a total of `r length(accel_df$day_id)` days, which was collected from a 63 year-old male admitted to Columbia's Center for Advanced Cardiac Care. For each day of observation it is noted what day of the week it was, including information from `r sum(accel_df$week_v_end)` weekdays. Finally, for each day there are 1440 activities reported, which account for each minute in the day, yielding `r format(35*1440, scientific = FALSE)` accelerometer observations for physical activity.

```{r accel_table}
accel_table <-
  accel_df %>% 
  group_by(day_id, week_v_end) %>% 
  select(-day, -week) %>% 
  mutate(sum = rowSums(across(activity_1:activity_1440))) %>% 
  select(day_id, week_v_end, sum) %>% 
  pivot_wider(names_from = "week_v_end", values_from = "sum") %>%
  knitr::kable(col.names = c("Observation Day (Day ID)", "Weekday","Weekend"),
               align = "ccc",
               digits = 0,
               caption = "Amount of activity per given observational day")

accel_table
```
When looking at the above table noting the amount of activity recorded for a given observational day, the most notable feature are the 2 weekends in which the patient had days recording the minimal amount of activity (i.e. a default of 1440, or 1 baseline activity measured every minute). It appears that the patient's activity is evenly distributed between weekends and weekdays but generally hovers between 300000 and 500000. 

```{r accel_plots}
accel_plot <-
  accel_df %>% 
  pivot_longer(cols = starts_with("activity_"), 
               names_prefix = "activity_",
               names_to = "minute", 
               values_to = "activity") %>% 
  mutate(minute = parse_number(minute)) %>% 
  ggplot(aes(y = activity, x = minute, color = day)) + 
  geom_line() + 
  labs(x = "Minute", y = "Activity") +
  scale_x_continuous(breaks = seq(from = 0, to = 1440, by = 240))

accel_plot

accel_facet <-
  accel_df %>% 
  pivot_longer(cols = starts_with("activity_"), 
               names_prefix = "activity_",
               names_to = "minute", 
               values_to = "activity") %>% 
  mutate(minute = parse_number(minute)) %>% 
  ggplot(aes(y = activity, x = minute, color = day)) + 
  geom_point() + 
  labs(x = "Minute", y = "Activity") +
  scale_x_continuous(breaks = seq(from = 0, to = 1440, by = 240)) +
  facet_grid(day ~ .) 

accel_facet
```

Based on the first graph it is not readily apparent if there are any patterns between the different days of the weeks (in which Monday is noted to be day 1 and Sunday day 7). However, once all of the days are separated it is noted that there are moments of more activity later in the day during the majority of the week while most activity on Sunday is limited to the middle of the day (bottom graph). Furthermore, there are moments of increased activity earlier in the day during the week but activity has a later start on Sunday.